Eliminating object prior-bias from sparse-projection tomographic reconstructions

### 2021-10-06 paper/post_tci.tex
Reorganized the section on computational complexity.  Also removed (latex commented) the irrelevant part of pre-computation and parallelization since the comment has already been made earlier (now appears as a footnote).
Added the note about the code base in changes.pdf also

However, reconstruction from reduced views has been made possible by
assuming the data to exhibit certain properties such as smoothness in
image space~\cite{Essam2015}, sparsity in gradient space~\cite{TV} or sparsity under
certain mathematical transforms~\cite{Donoho,introCS} such as the
Discrete Cosine Transform (DCT).  There are multiple such
regularization priors such as the Compressed
Sensing based l1-ls prior~\cite{my_dicta_paper}. For the purposes of
demonstrating the efficacy of our method, we use the well known total Variation (TV)
regularization prior.
If $\boldsymbol{\mathcal{R}}$ represents the acquisition
model, then the TV-regularized solution is described by one that
minimizes a combination of least squares error and the $TV$-norm of
the object $\boldsymbol{x}$. Mathematically, we choose to minimize
$J_{TV}(\boldsymbol{x})$ where

@@

  \item From $\boldsymbol{y_{Q_i}}$, perform reconstructions of the
    template $Q_i$ using the $M$ different algorithms, for each of the
    $L$ previously scanned objects. Let this set be denoted by $Y
    \triangleq \{\{Y_{i}^j\}_{j=1}^M\}_{i=1}^L$ where $Y^{1}_i =
    Y^{\text{fdk}}_i$, $\forall i \in \{1,..,,L\}$. If the set of projection views for every test is fixed a priori, this set of low quality reconstructions can be produced offline and stored, in order to save computational costs. If new sets of projection views for every test are to be allowed, the low quality reconstructions can still be performed efficiently using parallelization.

@@

Most of the computation time in our technique is used for the
generation of low quality eigen-space by reconstructing each of the
templates using TV and FDK. If the set of projection views for every
test is fixed a priori, a reasonable assumption,
the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. With
this assumption, Table~\ref{table:compute_time} shows the computation
times for various stages of reconstruction of a 2D slice of Liver
(refer code in~\cite{codeRepo}) on MATLAB 2019a on an AMD 2920X
12-Core Processor machine with 64GB RAM. If however, new sets of
projection views for every test are to be allowed, the low quality
reconstructions can still be performed efficiently using
parallelization since these reconstructions are completely
independent. Here below, we describe the time complexity of
spatially-varying prior technique.  Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We note that in practice $L$ is independent of $N$ and is likely to be
$O(1)$  (e.g. about 12 scans in a year for disease prognosis).
Likewise $M$ is also $O(1)$. The speed of the iterative optimization is dependent on the number of imaging views, and is profoundly impacted by the actual numerical optimization technique used.  
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspace, which is, in practice linear, i.e., $O(N)$
\item $O(NLM)$ for deriving weights map, which is, in practice linear, i.e., $O(N)$
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper
viz. generating weights map is linear in the number of voxels,
although admittedly the burden $F$ of the alternate minimization procedure drives up
the cost.

@@

\begin{table}[!h]
 \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver (refer code in~\cite{codeRepo}) on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
 \begin{tabular}[c]{@{}l@{}} FBP (Sufficient \\ views)\end{tabular}             & \begin{tabular}[c]{@{}l@{}} TV (Moderate \\ views)  \end{tabular}                      & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map \\ (Very few \\views)\end{tabular} & \begin{tabular}[c]{@{}l@{}}Alternate\\ Minimization \\ (Very few \\views)\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.01} & \multicolumn{1}{c|}{5.63} & \multicolumn{1}{c|}{0.01}                                             & \multicolumn{1}{c|}{34.00}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.

>>>>

However, reconstruction from reduced views has been made possible by
assuming the data to exhibit certain properties such as smoothness in
image space~\cite{Essam2015}, sparsity in gradient space~\cite{TV} or sparsity under
certain mathematical transforms~\cite{Donoho,introCS} such as the
Discrete Cosine Transform (DCT).  There are multiple such
regularization priors such as the Compressed
Sensing based l1-ls prior~\cite{my_dicta_paper}. For the purposes of
demonstrating the efficacy of our method, we use the well known Total Variation (TV)
regularization prior.
If $\boldsymbol{\mathcal{R}}$ represents the acquisition
model, then the TV-regularized solution is described by one that
minimizes a combination of least squares error and the $TV$-norm of
the object $\boldsymbol{x}$. Mathematically, we choose to minimize
$J_{TV}(\boldsymbol{x})$ where

@@

  \item From $\boldsymbol{y_{Q_i}}$, perform reconstructions of the
    template $Q_i$ using the $M$ different algorithms, for each of the
    $L$ previously scanned objects. Let this set\footnote{If the set of projection views for every test is fixed a priori, this set of low quality reconstructions can be produced offline and stored, in order to save computational costs. If new sets of projection views for every test are to be allowed, the low quality reconstructions can still be performed efficiently using parallelization.} be denoted by $Y
    \triangleq \{\{Y_{i}^j\}_{j=1}^M\}_{i=1}^L$ where $Y^{1}_i =
    Y^{\text{fdk}}_i$, $\forall i \in \{1,..,,L\}$.

@@

We discuss the computing burden of our method in two parts: the big-Oh
(symbolic) complexity, and the actual time taken on a specific dataset
on a specific computer. In the latter case, we provide the actual
code~\cite{codeRepo} for the purposes of reproduction.

\subsubsection{Symbolic}
Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We note that in practice $L$ is independent of $N$ and is likely to be
$O(1)$ (e.g., about 12 scans in a year for disease prognosis).
Likewise $M$ is also $O(1)$. The speed of the iterative optimization
is dependent on the number of imaging views, and is profoundly
impacted by the actual numerical optimization technique used.
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspaces, which is, in practice linear, i.e., $O(N)$.
\item $O(NLM)$ for deriving weights map, which is, in practice linear, i.e., $O(N)$.
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper viz.,
generating weights map is linear in the number of voxels, although
admittedly the burden $F$ of the alternate minimization procedure
drives up the cost (as can be seen in Table~\ref{table:compute_time}).

@@

\subsubsection{Example Case}
If the set of projection views for every test is fixed a priori, a
reasonable assumption, the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. 
% Most of the computation time in our technique is used for the
% generation of low quality eigen-space by reconstructing each of the
% templates using TV and FDK. 
With this as the basis, Table~\ref{table:compute_time} shows the
computation times for various stages of reconstruction of a 2D slice
of Liver on MATLAB 2019a on an AMD
2920X 12-Core Processor machine with 64GB RAM. 
% If however, new sets of
% projection views for every test are to be allowed, the low quality
% reconstructions can still be performed efficiently using
% parallelization since these reconstructions are completely
% independent. 
\begin{table}[!h]
 \centering
  \caption{Computation times (in seconds) for various stages of
    reconstruction of a 2D slice of Liver on MATLAB 2019a on an  AMD
    2920X 12-Core Processor machine with 64GB RAM. Our code is
    available for inspection~\cite{codeRepo}.}
\begin{tabular}{|l|l|l|l|}
\hline
 \begin{tabular}[c]{@{}l@{}} FBP (Sufficient \\ views)\end{tabular}             & \begin{tabular}[c]{@{}l@{}} TV (Moderate \\ views)  \end{tabular}                      & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map \\ (Very few \\views)\end{tabular} & \begin{tabular}[c]{@{}l@{}}Alternate\\ Minimization \\ (Very few \\views)\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.01} & \multicolumn{1}{c|}{5.63} & \multicolumn{1}{c|}{0.01}                                             & \multicolumn{1}{c|}{34.00}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.
